---
title: "Homework 04: Tidy data and joins"
author: "RH"
date: "October 8, 2017"
output:  
  github_document: 
    toc: true
    toc_depth: 5


---


The goal of this homework is to solidify your data wrangling skills by working some realistic problems in the grey area between data aggregation and data reshaping.
You can find the instructions for this assignment [here](http://stat545.com/hw04_tidy-data-joins.html)

```{r setup, include=FALSE}
## Loading libraries
library(tidyverse)
library(gapminder)
library(knitr)

```


# Choose your own adventure:

## Data reshaping

Activity #2

Make a tibble with one row per year and columns for life expectancy for two or more countries.
Use knitr::kable() to make this table look pretty in your rendered homework.
Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.


```{r}
#first let's create a new dataset with our 2 countries, Bangladesh and Canada
bdcan <- gapminder %>%
      filter(country %in% c("Bangladesh", "Canada")) %>%
      select(year, country, lifeExp) %>%             # selecting only interested values
      spread(key = "country", value = "lifeExp") 
      
#let's have a look at this data, we should have 24 observations
knitr::kable(bdcan, format = "markdown", digits = 2)
```

```{r}
# and now let's plot the data 
 ggplot(bdcan, aes(x = year, legend = Bangladesh)) + 
  geom_point(aes(y = Bangladesh), colour = "dark blue") + 
    geom_text(data = filter(bdcan, year == 2007), 
        aes(x= year, y= Bangladesh, label = "Bangladesh"), 
        colour = "dark blue",
        nudge_x = -3, nudge_y = 2) + 
  geom_point(aes (y = Canada), colour = "red") + 
    geom_text(data = filter(bdcan, year == 2007), 
        aes(x= year, y= Canada, label = "Canada"), 
        colour = "red",
        nudge_x = -3, nudge_y = 2) + 

  
  labs(title = "Life Expectancy of Bangladesh and Canada",
       x = "year", 
       y = "Life expectancy") + 
  theme_minimal()
```



## Data Joining

Activity #1

Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a  dplyr join function and make some observations about the process and result. Explore the different types of joins. Examples of a second data frame you could build:
One row per country, a country variable and one or more variables with extra info, such as language spoken, NATO membership, national animal, or capitol city. If you really want to be helpful, you could attempt to make a pull request to resolve this issue, where I would like to bring ISO country codes into the gapminder package.
One row per continent, a continent variable and one or more variables with extra info, such as northern versus southern hemisphere.
