---
title: "hw02"
author: "RH"
date: "September 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(gapminder)
library(tidyverse)
```

## Exploring the gapminder object

What type of data is it?

```{r}
data.class(gapminder)
summary (gapminder)

```


## Exploring individual variables

- categorical variable: **country**
- quantitative variable: **pop**

What are possible values (or range, whichever is appropriate) of each variable?
```{r}
summary(gapminder$pop) 
str(gapminder$country)

```

What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.
Feel free to use summary stats, tables, figures. We’re NOT expecting high production value (yet).

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Exploring plot types


## Use filter(), select() and %>%


```{r}
filter(gapminder, country == c("Rwanda", "Afghanistan"))
```
## More

Evaluate this code and describe the result. Presumably the analyst’s intent was to get the data for Rwanda and Afghanistan. Did they succeed? Why or why not? If not, what is the correct way to do this?

