---
title: 'Homework 06: Data wrangling wrap up'
author: "RH"
date: "November 8, 2017"
output:  
  html_document: 
    toc: true
    toc_depth: 5
---


```{r setup, include=FALSE}
## Loading libraries
library(tidyverse)
library(gapminder)
library(knitr)
library(stringr)

```

Pick (at least) two of the six (numbered) topics below and do one of the exercise prompts listed, or something comparable using your dataset of choice.


I pick topic 1 & 2


### 1. Character data

Read and work the exercises in the Strings chapter or R for Data Science - http://r4ds.had.co.nz/strings.html


##### 14.2.5 Exercises

In code that doesn’t use stringr, you’ll often see paste() and paste0(). What’s the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of NA?

paste0(..., collapse) is equivalent to paste(..., sep = "", collapse), slightly more efficiently. However, paste() coerces NA_character_, the character missing value, to "NA" which is sometimes useful and sometimes not. Also if we pass several vectors to paste0, they are concatenated in a vectorized way.

```{r}
x <- c("A", 1, NA, "B")
paste(x, collapse = "" )
paste0(x, collapse = "" )
```


In your own words, describe the difference between the sep and collapse arguments to str_c().
- The separator argument, sep, gives the character used to separate the different object by putting this character between inputs
- collapse is used to put together input vectors into a string

Use str_length() and str_sub() to extract the middle character from a string. What will you do if the string has an even number of characters?

```{r}
y <- "hellooworld"
str_sub(y, floor(str_length(y)/2 ), floor(str_length(y)/2 ))
```


What does str_wrap() do? When might you want to use it?

- it is a character vector of re-wrapped strings.
- it's useful if you need to wrap sentences based on screen sizes, for instance.

What does str_trim() do? What’s the opposite of str_trim()?
- it trims whitespaces from the  start and end of string. The opposite to str_trim() function is str_pad().


##### 14.3.1.1 Exercises

Explain why each of these strings don’t match a \: "\", "\\", "\\\".
- in order to match a "\" you need an escape which is "\\". But to match that you need to escape it first - i.e To create that regular expression, you need to use a string, which also needs to escape \. That means to match a literal \ you need to write "\\\\".

How would you match the sequence "'\?
- "'\\\\"

What patterns will the regular expression \..\..\.. match? How would you represent it as a string?
- it would match "......"
- representing it as a string "\\\\.\.\\\\.\.\\\\.\."


##### 14.3.2.1 Exercises

How would you match the literal string "$^$"?
- "\$\^\$"

Given the corpus of common words in stringr::words, create regular expressions that find all words that:

Start with “y”.
End with “x”
Are exactly three letters long. (Don’t cheat by using str_length()!)
Have seven letters or more.
Since this list is long, you might want to use the match argument to str_view() to show only the matching or non-matching words.
```{r}
str_view(stringr::words, "^y", match = TRUE)
str_view(stringr::words, "x$", match = TRUE)
str_view(stringr::words, ".......", match = TRUE)
```

##### 14.3.3.1 Exercises

Create regular expressions to find all words that:

Start with a vowel.
- "^(a|o|e|i|u)"

That only contain consonants. (Hint: thinking about matching “not”-vowels.)
- "[^a|o|e|i|u]"

End with ing or ise.
- "(ing$)|(ise$)"

##### 14.3.4.1 Exercises

Describe the equivalents of ?, +, * in {m,n} form.
?: 0 or 1			{0,1}
+: 1 or more		{1,}
*: 0 or more		{0,}


Create regular expressions to find all words that:

Start with three consonants. - "^([^a|o|e|i|u])++"
Have three or more vowels in a row. - "(a|o|e|i|u)++"


##### 14.3.5.1 Exercises

1. Describe, in words, what these expressions will match:

(.)\1\1						- same letter repeated 3 times
"(.)(.)\\2\\1"				- 
(..)\1 						- 
"(.).\\1.\\1"
"(.)(.)(.).*\\3\\2\\1"

2. Construct regular expressions to match words that:

Start and end with the same character.

Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)

Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)







### 2. Writing functions